\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-10 paper/main.tex
.

Other biases: evaluate on balanced data, not all

\todo[inline]{Likely discussion:}
The aim of \name is to predict candidate drugs that target a given
protein; the challenge is to develop a training and evaluation scheme
that does not simply overfit to the inherent biases in training and
testing data.
In general, when performing cross-validation for DTI prediction, the
options are to split over 
\begin{enumerate}
	\item split over drugs,
	\item split over drug--target pairs, or
	\item split over proteins
\end{enumerate}
where the first and third option concern splitting drugs and proteins,
respectively, into train, validation and test sets, and arranging the
corresponding drug-target interactions. They ensure that at least
parts of the interactions are not seen during training and evaluate
either how well targets are predicted for unseen drugs or unseen
proteins. Hereby, different training and prediction schemes lead to
divergent expressiveness of the resulting model.

A naive predictor (ranking proteins) and predict each drug similarly
achieves cutting edge performance ($87.5$ AUROC for whole dataset,
$85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
prot focused microAUC possible. $\rightarrow$ hub proteins

Yamanishi Dataset is only partially suitable for comparing results, if
everybody just derives a suitable subset (DTIGEMS)

This also applies to drug target affinity prediction. We were hereby
able to roughly reproduce the results from MolTrans (Bioinformatics)
on BioSnap

%\subsection{Features to use in predicting drug targets}

\todo[inline]{Discussion:}
The lack of stratification, only impacts the not
considered area under precision recall curve (AUPRC) and not the Macro
AUROC score, while also supporting the expressiveness of $MicroAUC_p$
with more data points.

@@

For the bottom-up approach we build a model that only relies on
molecular features, which we will discuss in more detail in the
following methods chapter. For the combination of both approaches we
now attach the predictions to the protein-protein interaction graph as
node features for future graph learning steps. In this graph we tried
to find both patterns and regions for each drug that could be of
interest through application of different graph convolutional layers,
which in return represent the feature for each protein. Representing
the drug we take the drug-drug interaction graph and the semantic
similarity over side effects which we will explain in the following
paragraphs.

Combining bottom-up and top-down features improves state of the art.

Evaluating DTI predictions is non-trivial due to numerous biases in
the training and evaluation datasets.

>>>>

% \todo[inline]{Likely discussion:}
% The aim of \name is to predict candidate drugs that target a given
% protein; the challenge is to develop a training and evaluation scheme
% that does not simply overfit to the inherent biases in training and
% testing data.
% In general, when performing cross-validation for DTI prediction, the
% options are to split over 
% \begin{enumerate}
% 	\item split over drugs,
% 	\item split over drug--target pairs, or
% 	\item split over proteins
% \end{enumerate}
% where the first and third option concern splitting drugs and proteins,
% respectively, into train, validation and test sets, and arranging the
% corresponding drug-target interactions. They ensure that at least
% parts of the interactions are not seen during training and evaluate
% either how well targets are predicted for unseen drugs or unseen
% proteins. Hereby, different training and prediction schemes lead to
% divergent expressiveness of the resulting model.

% A naive predictor (ranking proteins) and predict each drug similarly
% achieves cutting edge performance ($87.5$ AUROC for whole dataset,
% $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
% prot focused microAUC possible. $\rightarrow$ hub proteins

% Yamanishi Dataset is only partially suitable for comparing results, if
% everybody just derives a suitable subset (DTIGEMS)

% This also applies to drug target affinity prediction. We were hereby
% able to roughly reproduce the results from MolTrans (Bioinformatics)
% on BioSnap

% %\subsection{Features to use in predicting drug targets}

%\todo[inline]{Discussion:}
% The lack of stratification, only impacts the not
% considered area under precision recall curve (AUPRC) and not the Macro
% AUROC score, while also supporting the expressiveness of $MicroAUC_p$
% with more data points.

@@

We developed \name{} as a machine learning model that combines
molecular features and functional information with an interaction
network using graph neural networks to predict drugs that may target
specific proteins. In this task, \name{} improves over several state
of the art methods. We demonstrated that functional and phenotypic
information localizes on the interaction network whereas molecular
information does not. Moreover, we showed that drug--target
interaction prediction datasets have some inherent biases that affect
the performance of models. This led us to conclude that drug--target
interaction prediction is not a single computational problem but a set
of multiple problems. Experimental evaluation of drug--target
interaction prediction methods must be carefully designed to reflect
the problem the model aims to solve, and the interpretation of
performance results should be aligned with the specific problem.
